\input{../header.tex}

\title{
	$\Sum \text{math}(y)$\\
	Rings
}
\author{Jonas Linßen}

\begin{document}
	\maketitle
	\tableofcontents

	\newpage
	\section{Rings}
	\subsection{Rings}

	\begin{definition}
		A \textbf{(unital) ring} is a set $R$ together with a \textbf{addition} $+: R \times R --> R$ and a \textbf{multiplication} $\cdot: R \times R --> R$ such that 
		\begin{enumerate}[$\bullet$]
			\item{
				$(R,+)$ is an \textit{abelian group}, i.e. $+$ is associative and commutative, any element $a$ has a unique additive inverse $-a$ and there is an additive identity element $0$. 
			}
			\item{
				$(R,\cdot)$ is a \textit{monoid}, i.e. $\cdot$ is associative and there is a multiplicative identity element $1$.
			}
			\item{
				the following \textit{distributive law} holds for any $a,b,c \in R$:
				\begin{align*}
					a \cdot (b+c) &= a \cdot b + a \cdot c
				\end{align*}
			}
		\end{enumerate}

		A ring $R$ is said to be \textbf{commutative}, if its multiplication is.

		If $0 = 1$ the ring necessessarily becomes $R = \{0\}$, which is called the \textbf{zero ring} or \textbf{trivial ring}.
	\end{definition}

	For sake of intuition we state the following examples, even if we will define them thoroughly in later sections.
	\begin{enumerate}[$\bullet$]
		\item{
			The integer numbers $\Z$, the rational numbers $\Q$, the real numbers $\R$ and the complex numbers $\C$ are commutative rings, equipped with their standard addition and multiplication.
		}
		\item{
			The set $\ncat{Set}(X,R)$ of functions from a set $X$ to a (commutative) ring $R$ is a (commutative) ring, when equipped with pointwise addition and pointwise multiplication. The identities are the corresponding constant maps.
		}
		\item{
			The set of endomorphisms $\End(A)$ of an abelian group, together with pointwise addition and composition as multiplication. In general it will be a non-commutative ring.
		}
	\end{enumerate}

	To compare different rings we need to have a definition of structure preserving map.

	\begin{definition}
		A \textbf{morphism of rings} is a function $f:R --> S$ between rings $R$ and $S$, which satisfies for all $a,b,c \in R$ that
		\begin{equation*}
			f(ab + c) = f(a)f(b) + f(c) \;\;\;\;\text{and}\;\;\;\; f(1_R) = 1_S.
		\end{equation*}
		As the identity function $1_R: R --> R$ is a morphism of rings for any ring $R$ and morphisms of rings are closed under composition, we get a category $\ncat{Ring}$. The full subcategory of commutative rings will be denoted $\ncat{CRing}$.
	\end{definition}

	\begin{lemma}[Properties of $\ncat{Ring}$ and $\ncat{CRing}$]
		In $\ncat{Ring}$ we have the following.
		\begin{enumerate}[(i)]
			\item{
				A morphism of rings $f:R-->S$ is a morphism of the underlying \textit{abelian groups} $(R,+)$ and $(S,+)$, i.e. there is a forgetful functor $\fun{U}:\ncat{CRing} --> \ncat{Ab}$. In particular it holds that
				\begin{equation*}
					f(0) = 0 \;\;\;\;\text{and}\;\;\;\; f(-a) = -f(a).
				\end{equation*}
			}
			\item{
				A morphism of rings $f:R --> S$ is mono if and only if it is injective. Using (i) this is equivalent to the equation $\ker f := \{r \in R \mid f(r) = 0\} = \{0\}$.
			}
			\item{
				The trivial ring $1$ is terminal and the ring of integers $\Z$ is initial, both in $\ncat{Ring}$ and $\ncat{CRing}$.
			}
			\clearpage\vspace{-2em}\item{
				Both $\ncat{Ring}$ and $\ncat{CRing}$ are complete categories and the forgetful functors $\fun{U}:\ncat{Ring} --> \ncat{Set}$ respectively $\fun{U}:\ncat{CRing} --> \ncat{Set}$ create limits. Concretely
				\begin{enumerate}[$\bullet$]
					\item{
						Products of rings $(R_i)_{i \in I}$ are given by componentwise ring structure on the product $\smash{\Product \limits_{i \in I} R_i}$ of the underlying sets.
					}\vspace{.5em}
					\item{
						The equalizer of morphisms $f_1,f_2:R --> S$ is the subring given by the equalizer of the underlying functions $\{r \in R \mid f_1(r) = f_2(r)\}$.
					}
				\end{enumerate}
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				By definition.
			}
			\item{
				Follows from the analogous statement for abelian groups.
			}
			\item{
				The first statement holds by definition. For the second note that $n = 1 + ... + 1$ in $\Z$ and use a morphism of rings has to preserve $1$.
			}
			\item{
				 In both cases it suffices to check that the unique morphism into $\Product \limits_{i \in I} R_i$, respectively $\eq(f_1,f_2)$ of the universal property in $\ncat{Set}$ is a morphism of rings.
			}
		\end{enumerate}\vspace{-2em}
	\end{sketch}

	In arbitrary rings, counterintuitive things may happen. Consider for example the binary product $R^2$ of a ring $R$ with itself. The two elements $(1,0)$ and $(0,1)$ are clearly nonzero, but satisfy $(1,0)\cdot(0,1) = (0,0)$. We will name elements of satisfying this or similar conditions in the following definition.

	\begin{definition}
		A \textbf{zero divisor} in a ring $R$ is an element $a$, such that there exists a nonzero element $b$ such that $ab = 0$. A \textbf{nilpotent element} $a$ satisfies that $a^n = 0$ for some $n \in \N$. In particular nilpotent elements are zero divisors.
		An element, which is not a zero divisor, is sometimes called \textbf{regular}. A \textbf{unit} in a ring $R$ is an element $a$, for which there exists a multiplicative inverse, i.e. an element $a^{-1}$ satisfying $aa^{-1} = 1 \;{\color{gray}(= a^{-1}a)}$.
		The subset of all units in $R$ will be denoted by $R^\times$. 

		A ring, in which $0$ it the only zero divisor, is called an \textbf{integral domain}.

		A ring, in which every element except $0$ is a unit, is a \textbf{field}.
	\end{definition}

	Note that $0$ is a zero divisor unless the ring is the trivial ring. In particular, the trivial ring is considered \underline{not} to be an integral domain.

	\begin{lemma}[Properties of Units and Zero Divisors]
		\vspace{-1.5em}\begin{enumerate}[(i)]
			\item{
				Let $a$ be a regular element. Then $ab=ab'$ implies $b=b'$. In particular this cancellation rule holds in any integral domain.
			}
			\item{
				No zero divisor is a unit and vice versa. In particular every field is an integral domain.
			}
			\item{
				Let $f:R --> S$ be a morphism of rings. If $a$ is a zero divisor / nilpotent / unit in $R$, then $f(a)$ is a zero divisor / nilpotent / unit in $S$.
			}
			\item{
				Let $f: K --> R$ be a morphism of rings, where $K$ is a field. Then $f$ is mono.
			}
			\item{
				For any nilpotent element $r$ in some ring $R$, the element $1-r$ is a unit.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				By assumption.
			}
			\item{
				Holds for the trivial ring by definition. In a nontrivial ring use contradiction.
			}
			\item{
				By definition.
			}
			\item{
				Follows from (ii) via $\ker f = \{0\}$.
			}
			\item{
				Given $r^n = 0$ the inverse of $1-r$ is given by $\smash{\Sum \limits_{k=0}^{n-1}r^k}$.
			}
		\end{enumerate}\vspace{-2em}
	\end{sketch}

	\newpage
	\subsection{Ideals and Factor Rings}
	\begin{definition}
		Let $R$ be a ring. An \textbf{ideal} $\ideal{a}$ is an \textit{abelian subgroup} $\ideal{a} \leq_\ncat{Ab} R$, which is closed under multiplication with $R$. We write $\ideal{a} \diveq R$ and say that it is \textbf{proper}, if $\ideal{a} \divneq R$.

		The \textbf{ideal generated by} a set of elements $X$ is given by 
		\begin{equation*}
			\Gen{X} = \Intersection \limits_{X \subseteq \ideal{a} \diveq R} \ideal{a} = \{\Sum \limits_{i=1}^n r_ix_i \mid n \in \N, x_i \in X, r_i \in R\}.
		\end{equation*}
	\end{definition}

	Given a subset $X \subseteq R$ and an ideal $\ideal{a}\diveq R$ define the ideal $(\ideal{a}\;:X) = \{r \in R \mid r\cdot X \subseteq \ideal{a}\}$. \TODO{name of this construction?}

	\begin{lemma}[Properties of Ideals]
		\vspace{-1.5em}\begin{enumerate}[(i)]
			\item{
				Let $R$ be a ring and $\ideal{a},\ideal{b} \diveq R$ ideals. Then
				\begin{align*}
					\ideal{ab} &= \Gen{ab \mid a \in \ideal{a}, b \in \ideal{b}}\\
					\ideal{a} \intersection \ideal{b} &= \{r \mid r \in \ideal{a}, r \in \ideal{b}\}\\
					\ideal{a} + \ideal{b} &= \{a + b \mid a \in \ideal{a}, b \in \ideal{b}\}
				\end{align*}
				are ideals and it holds that 
				\begin{equation*}
					\ideal{ab} \subseteq \ideal{a} \intersection \ideal{b} \subseteq \ideal{a}, \ideal{b} \subseteq \ideal{a} + \ideal{b}.
				\end{equation*}
				If $\ideal{a}$ and $\ideal{b}$ are \textit{comaximal}, i.e. $\ideal{a} + \ideal{b} = R$, the equation $\ideal{ab} = \ideal{a} \intersection \ideal{b}$ holds.
			}
			\item{
				Let $f:R --> S$ be a morphism of rings and $\ideal{b} \diveq S$ be an ideal in $S$.

				Then $\ker f \diveq R$ and $f^{-1}(\ideal{b}) \diveq R$ are ideals in $R$.
			}
			\item{
				Let $f:R -->> S$ be an epimorphism of rings and $\ideal{a} \diveq R$ be an ideal in $R$.

				Then $f(\ideal{a}) \diveq S$ is an ideal in $S$.
			}
			\item{
				If an ideal $\ideal{a} \diveq R$ in a ring $R$ contains a unit, it satisfies $\ideal{a}=R$. In particular $R$ is a field if and only if it has no proper ideals besides $\Gen{0}$.
			}
			\item{
				Given rings $R_1,R_2$ any ideal in $R_1 \times R_2$ is of the form $\ideal{a}_1 \times \ideal{a}_2$, where $\ideal{a}_i$ is an ideal in $R_i$. \TODO{arbitrary products counterexample in wikipedia product ring}
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				The ideal property and inclusions hold by definition. For the equation assuming comaximality note that $1 = a_0+b_0$ for some $a_0 \in \ideal{a}, b_0 \in \ideal{b}$. Given $r \in \ideal{a}\intersection\ideal{b}$ we then have $r = r\cdot 1 = r\cdot(a_0+b_0) = a_0r + rb_0 \in \ideal{ab}$.
			}
		\end{enumerate}
		\hspace{.15cm}(ii) -- (v) By definition.
	\end{sketch}

	\begin{definition}
		Let $R$ be a ring and $\ideal{a} \diveq R$ be an ideal in $R$. As $\ideal{a} \leq_\ncat{Ab} R$ is an abelian subgroup we can construct the factor group $R/\ideal{a}$, with addition given by
		\begin{equation*}
			(x + \ideal{a}) + (y + \ideal{a}) = (x+y) + \ideal{a}.
		\end{equation*}
		Furthermore, since $\ideal{a}$ is closed under multiplication with $R$, the canonic multiplication
		\begin{equation*}
			(x + \ideal{a}) \cdot (y + \ideal{a}) = (x \cdot y) + \ideal{a}
		\end{equation*}
		is well defined, turning $R / \ideal{a}$ into a \textbf{factor ring} (or \textbf{quotient ring} or \textbf{residue class ring}). This construction yields a canonic \textbf{projection morphism} $\pi_\ideal{a}: R -->> R/\ideal{a}$, which is epi and satisfies $\ker \pi_\ideal{a} = \ideal{a}$.
	\end{definition}

	\begin{lemma}[Properties of the Factor Ring]
		Let $R$ be a ring and $\ideal{a} \diveq R$ be an ideal.
		\begin{enumerate}[(i)]
			\item{
				Given parallel morphisms of rings $f_1,f_2:R-->S$ define $\ideal{a} = \Gen{s \in S \mid s=f_1(r)=f_2(r) \text{ for some }r \in R}_S$. Then\vspace{-1em}
				\begin{equation*}
					S/\ideal{a}=\coeq(
					\begin{diagram}
						\twobyone{R}{S}
						\arrow[higher]{w}{e}{f_1}[above]
						\arrow[lower]{w}{e}{f_2}[below]
					\end{diagram}
					)
				\end{equation*}
			}
		\end{enumerate}
	\end{lemma}

	\begin{theorem}[Isomorphism Theorem]
		Let $f:R --> S$ be a morphism of rings such that $\ideal{a} \diveq R$ is an ideal in $R$, which satisfies $\ideal{a} \subseteq \ker f$. Then the following universal property holds.
		\begin{equation*}
			\begin{diagram}
				\threebytwo[small]
					{R}{}{S}
					{}{R/\ideal{a}}{}

				\arrow{nw}{ne}{f}[above]
				\arrow[epi]{nw}{s}{\pi_\ideal{a}}[below left]
				\arrow[dashed]{s}{ne}{\exists!\tilde{f}}[below right]
			\end{diagram}
		\end{equation*}
		In particular, for $\ideal{a} = \ker{f}$ we get a epi-mono-factorization (i.e. $\tilde f$ is mono) and 
		\begin{equation*}
			R / {\ker f} \isom \Im f.
		\end{equation*}
	\end{theorem}
	\begin{sketch}
		Check that 
		\begin{equation*}
			\tilde{f}:
			\begin{array}{rrcl}
					R/\ideal{a}	&\longrightarrow	&S\\
					x+\ideal{a}	&\longmapsto 		&f(x)
			\end{array}
		\end{equation*}
		is a well defined morphism of rings. Uniqueness follows from $\pi$ being epi.
	\end{sketch}

	\begin{theorem}[Chinese Remainder Theorem]
		Let $R$ be a ring and $\ideal{a}_1, ..., \ideal{a}_n \diveq R$ be a ideals in $R$ such that $\ideal{a}_i + \ideal{a}_j = R$ for all $i \neq j$. Then
		\begin{equation*}
			R / {\Intersection \limits_{i=1...n} \ideal{a}_i} \isom R/\ideal{a}_1 \times ... \times R/\ideal{a}_n.
		\end{equation*}
	\end{theorem}
	\begin{sketch}
		The morphism
		\begin{equation*}
			f:
			\begin{array}{rrcl}
					R	&\longrightarrow	&R/\ideal{a}_1 \times ... \times R/\ideal{a}_n\\
					r	&\longmapsto 		&(r+\ideal{a}_1, ..., r + \ideal{a}_n)
			\end{array}
		\end{equation*}
		satisfies $\Intersection \limits_{i=1...n} \ideal{a}_i = \ker f$ and thus by the isomorphism theorem there is an injective morphism of rings in place of the claimed isomorphism.

		For surjectivity consider $(r_1 + \ideal{a}_1, ..., r_n + \ideal{a}_n) = e_1f(r_1) + ... + e_nf(r_n)$, where $e_i$ denotes the \textit{standard unit vector} $(0,...,1+\ideal{a}_i,...,0)$. Suppose for all $i$ there is $c_i \in R$ with $e_i = f(c_i)$. Then
		\begin{align*}
			f(c_1r_1 + ... + c_nr_n) &= f(c_1)f(r_1) + ... + f(c_n)f(r_n)\\
			&=e_1f(r_1) + ... + e_nf(r_n)\\
			&=(r_1 + \ideal{a}_1, ..., r_n + \ideal{a}_n).
		\end{align*}
		By pairwise comaximality we have $a_{ji} \in \ideal{a}_i$ and $a_{ij} \in \ideal{a}_j$ with $a_{ij} + a_{ji} = 1$ for all $i\neq j$. Setting $c_i = \Product\limits_{j\neq i}a_{ij}$ we obtain $c_i \in \ideal{a}_j$, i.e. $c_i + \ideal{a}_j = 0 + \ideal{a}_j$, for all $j \neq i$ and 
		\begin{equation*}
			c_i + \ideal{a}_i = \Product \limits_{j\neq i}(a_{ij}+\ideal{a}_i) = \Product \limits_{j \neq i}(a_{ij} + a_{ji} + \ideal{a}_i) = \Product \limits_{j \neq i}(1 + \ideal{a}_i) = 1 + \ideal{a}_i.\vspace{-2.8em}
		\end{equation*}
	\end{sketch}

	\begin{definition}
		Let $R$ be a ring. A \textbf{prime ideal} is a proper ideal $\ideal{p} \divneq R$, which satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				The factor ring $R/\ideal{p}$ is an integral domain.
			}
			\item{
				It holds that $ab \in \ideal{p}$ implies that $a \in \ideal{p}$ or $b \in \ideal{p}$.
			}
			\item{
				The set $R \setminus \ideal{p}$ is a multiplicative system, i.e. contains $1$ and is closed under multiplication.
			}
			\item{
				For ideals $\ideal{a},\ideal{b} \diveq R$ it holds that $\ideal{a}\ideal{b} \subseteq \ideal{p}$ implies $\ideal{a} \subseteq \ideal{p}$ or $\ideal{b} \subseteq \ideal{p}$.
			}
		\end{enumerate}
		The set of all prime ideals is called the \textbf{spectrum} of $R$ and is denoted by $\Spec R$.
	\end{definition}
	\begin{sketch}
		(i)$<==>$(ii)$<==>$(iii) just reformulations.\\
		(iv)$==>$(ii) consider the principal ideals $\Gen{a},\Gen{b}$.\\
		\TODO{???$==>$(iv)}
	\end{sketch}

	\begin{lemma}[Properties of Prime Ideals]\vspace{-1.5em}
		\begin{enumerate}[(i)]
			\item{
				A ring $R$ is an integral domain if and only if $\Gen{0}$ is a prime ideal.
			}
			\item{
				Let $f:R-->S$ be a morphism of rings and $\ideal{q} \diveq S$ a prime ideal in $S$. Then $f^{-1}(\ideal{q})$ is a prime ideal in $R$.
			}
			\item{
				Any prime ideal contains a minimal prime ideal, i.e. one which is minimal with respect to inclusion.
			}
			\item{
				Let $\ideal{p}$ be a prime ideal in a ring $R$. For ideals $\ideal{a},\ideal{b} \diveq R$ it holds that $\ideal{a} \intersection \ideal{b} \subseteq \ideal{p}$ implies $\ideal{a} \subseteq \ideal{p}$ or $\ideal{b} \subseteq \ideal{p}$.
			}
			\item{
				The prime ideals in a product of rings $R_1 \times R_2$ are of the form $\ideal{p}_1 \times R_2$ or $R_1 \times \ideal{p}_2$, where $\ideal{p}_i$ is a prime ideal in $R_i$.

				\TODO{This does not hold for arbitrary products, see https://math.stackexchange.com/questions/1049867/prime-ideals-in-an-arbitrary-direct-product-of-rings.}
				%The prime ideals in a product of rings $\smash{\Product \limits_{i\in I} R_i}$ are of the form $\smash{\ideal{p}_i \times \Product \limits_{j \neq i} R_j}$, where $\ideal{p}_i$ is a prime ideal in $R_i$.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				By definition.
			}
			\item{
				By definition.
			}
			\item{
				Use Zorn on the set $\{S \mid S \neq R \text{ multiplicative system with }R \setminus \ideal{p} \subseteq S\}$.
			}
			\item{
				By $\ideal{ab} \subseteq \ideal{a} \intersection \ideal{b}$.
			}
			\item{
				Note that for $(R_1 \times R_2) / (\ideal{a}_1 \times \ideal{a}_2)$ to be an integral domain, at most one of the $\ideal{a}_i$ can be a proper ideal.
			}
			\vspace{-1.8em}
		\end{enumerate}
	\end{sketch}

	\begin{definition}
		Let $R$ be a ring and $\mathcal{F}$ be a family of ideals with $R \in \mathcal{F}$.

		$\mathcal{F}$ is an \textbf{Oka family}, if for any ideal $\ideal{a} \diveq R$ and element $x \in R$ it holds that $\ideal{a} + \Gen{x} \in \mathcal{F}$ and $(\ideal{a}\;:x) \in \mathcal{F}$ implies $\ideal{a} \in \mathcal{F}$.

		$\mathcal{F}$ is an \textbf{Ako family}, if for any ideal $\ideal{a} \diveq R$ and elements $x,y \in R$ it holds that $\ideal{a} + \Gen{x} \in \mathcal{F}$ and $\ideal{a} + \Gen{y} \in \mathcal{F}$ implies $\ideal{a} + \Gen{xy} \in \mathcal{F}$.
	\end{definition}

	\begin{lemma}[Lam Reyes Prime Ideal Principle]
		Let $R$ be a ring and $\mathcal{F}$ be a family of ideals with $R \in \mathcal{F}$. Denote by $\mathcal{F}'$ the set of ideals not in $\mathcal{F}$ and by $\Max \mathcal{F}'$ the set of maximal elements in $\mathcal{F}'$ with respect to inclusion.

		If $\mathcal{F}$ is Oka or Ako family, then $\Max \mathcal{F}' \subseteq \Spec R$.
	\end{lemma}
	\begin{proof}
		By contraposition.

		Suppose $\ideal{a} \in \Max\mathcal{F}'$ is not prime. Since $\ideal{a} \notin \mathcal{F}$ we have $\ideal{a} \neq R$, so there are $x,y \in R\setminus \ideal{a}$ with $xy \in \ideal{a}$. Then $\ideal{a} \subsetneq \ideal{a} + \Gen{x}, \ideal{a}+\Gen{y}, (\ideal{a}\;:x)$, since the former contains $a$ and the latter contain $y$. Hence $\ideal{a} + \Gen{x}, \ideal{a}+\Gen{y}, (\ideal{a}\;:x)$ belong to $\mathcal{F}$. But $\ideal{a}+\Gen{ab} = \ideal{a} \notin \mathcal{F}$, so $\mathcal{F}$ is neither Oka nor Ako.
	\end{proof}

	\begin{definition}
		Let $R$ be a ring. A \textbf{maximal ideal} is a proper ideal $\ideal{m} \divneq R$, which satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				The factor ring $R/\ideal{m}$ is a field.
			}
			\item{
				The ideal $\ideal{m}$ is maximal in $\{\ideal{a} \divneq R\}$ with respect to inclusion.
			}
			\item{
				The ideal $\ideal{m}$ is maximal in $\{\ideal{p} \divneq R \mid \ideal{p} \text{ prime}\}$ with respect to inclusion.
			}
		\end{enumerate}
		By definition any maximal ideal is prime. The set of all maximal ideals is denoted by $\mSpec R$.
	\end{definition}
	\begin{sketch}
		(i)$<==>$(ii) by definition

		(ii)$<==>$(iii) note that $\{R\}$ is an Ako family.
	\end{sketch}

	\begin{lemma}[Properties of Maximal Ideals]
		Any ideal is contained in a maximal ideal.
	\end{lemma}
	\begin{sketch}
		Use Zorn on the set of proper ideals containing the ideal.
	\end{sketch}

	\begin{definition}
		Let $R$ be a ring and $\ideal{a} \diveq R$ be an ideal. 

		The \textbf{radical} of $\ideal{a}$ is the ideal 
		\begin{equation*}
			\sqrt{\ideal{a}} = \{x \in R \mid \exists n \in \N: x^n \in \ideal{a}\} = \Intersection \limits_{\ideal{a} \subseteq \ideal{p} \in \Spec R} \ideal{p}.
		\end{equation*}
		An ideal $\ideal{a}$, which satisfies $\ideal{a} = \sqrt{\ideal{a}}$ is said to be \textbf{radical}.

		The \textbf{nilradical} of $R$ is the radical ideal
		\begin{align*}
			\nrad R &= \{x \in R \mid \exists n \in \N: x^n = 0\} = \Intersection \limits_{\ideal{p} \in \Spec R} \ideal{p}\\
			& \textcolor{red}{\overset{?}{=} \{x \in R \mid \exists n \in \N: 1 - x^n \in R^\times\}}.
		\end{align*}

		The \textbf{Jacobson radical} of $R$ is the radical ideal
		\begin{equation*}
			\jrad R = \Intersection \limits_{\ideal{m} \in \mSpec R} \ideal{m} = \{x \in R \mid \forall r \in R: 1 - rx \in R^\times\}.
		\end{equation*}
	\end{definition}

	\begin{lemma}[Properties of Radical Ideals]
		Let $R$ be a ring and $\ideal{a},\ideal{b} \diveq R$ be two ideals in R.
		\begin{enumerate}[(i)]
			\item{
				For any $n \in \N$ we have $\sqrt{\sqrt{\ideal{a}}} = \sqrt{\ideal{a}} = \sqrt{\ideal{a}^n}$.
			}
			\item{
				It holds that $\ideal{a} \subseteq \sqrt{\ideal{a}}$ and that $\ideal{a} \subseteq \ideal{b}$ implies $\sqrt{\ideal{a}} \subseteq \sqrt{\ideal{b}}$.
			}
			\item{
				We have $\sqrt{\ideal{a} \intersection \ideal{b}} = \sqrt{\ideal{a}} \intersection \sqrt{\ideal{b}}$
			}
			\item{
				The equation $\sqrt{\ideal{a}} = \sqrt{\ideal{b}}$ is equivalent to having $n,m \in \N$ such that $\ideal{a}^m \subseteq \sqrt{\ideal{b}}$ and $\ideal{b}^n \subseteq \sqrt{\ideal{a}}$.
			}
			\item{
				Any prime ideal $\ideal{p}$ is radical, i.e. satisfies $\sqrt{\ideal{p}} = \ideal{p}$.
			}
		\end{enumerate}
	\end{lemma}

	\begin{theorem}[Correspondence Theorem]
		Let $R$ be a ring and $\ideal{a} \diveq R$ be an ideal. The canonic projection $\pi: R -->> R/\ideal{a}$ exhibits inclusion preserving bijections
		\begin{equation*}
			\begin{array}{rcl}
				\{\text{ideals containing }\ideal{a}\} &\overset{\sim}{\longrightarrow} &\{\text{ideals in } R/\ideal{a}\}\\
				\{\text{radical ideals containing }\ideal{a}\} &\overset{\sim}{\longrightarrow} &\{\text{radical ideals in } R/\ideal{a}\}\\
				\{\text{prime ideals containing }\ideal{a}\} &\overset{\sim}{\longrightarrow} &\{\text{prime ideals in } R/\ideal{a}\}\\
				\{\text{maximal ideals containing }\ideal{a}\} &\overset{\sim}{\longrightarrow} &\{\text{maximal ideals in } R/\ideal{a}\}\\\\
				\ideal{b} &\longmapsto &\pi_\ideal{a}(\ideal{b})\\
				\pi_\ideal{a}^{-1}(\ideal{b}) &\reflectbox{$\longmapsto$} &\ideal{b}
			\end{array}
		\end{equation*}
	\end{theorem}
	\begin{sketch}
		The correspondence for ideals can be checked by composing the mutually inverse maps. That for maximality follows from the former preserving inclusion. The correspondences for prime ideals may be checked by considering elements. The one for radical ideals the follows from the fact that for surjective functions preimages and images respect intersections. 
	\end{sketch}

	\TODO{$R/\nrad(R)$ is free reduced ring}

	\newpage
	\subsection{Localization and Local Rings}

	\begin{definition}
		Let $S$ be a multiplicative system in a ring $R$, i.e. contain $1$ and be closed under multiplication. The relation
		\begin{align*}
			(r_1,s_1) \sim (r_2,s_2) <==> \exists f\in S: fr_1s_2 = fr_2s_1
		\end{align*}
		defines an equivalence relation on $R \times S$. We write $\dfrac{r}{s}$ for the equivalence class of $[(r,s)]$. By setting
		\begin{equation*}
			\dfrac{r_1}{s_1} + \dfrac{r_2}{s_2} = \dfrac{r_1s_2 + r_2s_1}{s_1s_2} \hspace{1cm}\text{and}\hspace{1cm} \dfrac{r_1}{s_1} \cdot \dfrac{r_2}{s_2} = \dfrac{r_1r_2}{s_1s_2}
		\end{equation*}
		we obtain a ring structure on $R \times S / \sim$. This is the so called \textbf{localization} $S^{-1}R$ of $R$ by $S$ and comes with the canonical map
		\begin{equation*}
			\begin{array}{rcl}
				\iota_S:R & \longrightarrow & S^{-1}R\\
				r & \longmapsto & \dfrac{r}{1}
			\end{array}.
		\end{equation*}
	\end{definition}

	\TODO{remark on factor $f$, field of fractions}

	\begin{lemma}[Properties of Localization]
		Let $S$ be a multiplicative system in a ring $R$.
		\begin{enumerate}[(i)]
			\item{
				The canonical map $\iota_S$ is injective, if and only if $S$ does not contain a zero divisor (in particular $S$ does not contain $0$).
			}
			\item{
				\begin{minipage}{\textwidth-4cm}
				Let $f:R-->R'$ be a morphism of rings and $S'$ be a multiplicative system in $R'$ with $f(S) \subseteq R'^\times$. Then there is a unique morphism $S^{-1}f: S^{-1}R --> R$ making the diagram on the right commute.
				\end{minipage}
				\begin{minipage}{4cm}
					\begin{equation*}
						\begin{diagram}
							\twobytwo
								{S^{-1}R}{R'}
								{R}{}
							\arrow{sw}{nw}{\iota_S}[left]
							\arrow{sw}{ne}{f}[below right]
							\arrow[dashed]{nw}{ne}{S^{-1}f}[above]
						\end{diagram}
					\end{equation*}
				\end{minipage}
			}
			\item{
				Given an ideal $\ideal{a} \diveq R$ denote $S/\ideal{a} := \pi_\ideal{a}(S)$ and $S^{-1}\ideal{a} := \Gen{\iota_S(\ideal{a})}_{S^{-1}R}$. Then there is a canonical isomorphism
				\begin{equation*}
					(S/\ideal{a})^{-1}(R/\ideal{a}) \isom S^{-1}R/S^{-1}\ideal{a}.
				\end{equation*}
			}
			\item{
				Given a family of ideals $(\ideal{a}_i)_{i \in I}$ in $R$ it holds that $S^{-1}(\Intersection \limits_{i \in I} \ideal{a}_i) = \Intersection \limits_{i \in I} (S^{-1}\ideal{a}_i)$.
			}
		\end{enumerate}
	\end{lemma}

	\begin{theorem}[Correspondence Theorem]
		Let $R$ be a ring and $S$ be a multiplicative system in $R$. The canonic morphism $\iota_S: R --> S^{-1}R$ induces an inclusion preserving injection
		\begin{equation*}
			\begin{array}{rcl}
				\{\text{ideals in } S^{-1}R\} & \longrightincl & \{\text{ideals }\ideal{a}\text{ with } \forall s \in S, r \in R \setminus \ideal{a}: sr \notin \ideal{a}\}\\
				\ideal{b} & \longmapsto & \iota_S^{-1}(\ideal{b})
			\end{array}
		\end{equation*}
		with section
		\begin{equation*}
			\hspace{-4.2cm}S^{-1}\ideal{a} \;\;\;\; \reflectbox{$\longmapsto$} \;\;\;\; \ideal{a},
		\end{equation*}
		which restricts to inclusion preserving bijections
		\begin{equation*}
			\begin{array}{rcl}
				\{\text{radical ideals in } S^{-1}R\} & \overset{\sim}{\longrightarrow} & \{\text{radical ideals }\ideal{r}\text{ with } \forall s \in S, r \in R \setminus \ideal{r}: sr \notin \ideal{r}\}\\
				\{\text{prime ideals in } S^{-1}R\} & \overset{\sim}{\longrightarrow} & \{\text{prime ideals }\ideal{p}\text{ with } \ideal{p} \intersection S = \emptyset\}\\
			\end{array}
		\end{equation*}
	\end{theorem}

	\begin{definition}
		A ring $R$ is said to be \textbf{local}, if it satisfies one of the following equivalent contitions.
		\begin{enumerate}[(i)]
			\item{
				$R$ has exactly one maximal ideal $\ideal{m}_R$.
			}
			\item{
				The complement $R \setminus R^\times$ is an ideal.
			}
			\item{
				$R$ is nontrivial and for any $r \in R$ it holds that $r$ or $1-r$ are invertible.
			}
		\end{enumerate}

		Recall that in general preimages of maximal ideals need not be maximal. A morphism $f:R-->R'$ between local rings is a \textbf{local morphism}, if it satisfies $f^{-1}(\ideal{m}_{R'}) = \ideal{m}_R$ or equivalently that $f(r)$ is invertible if and only if $r$ is invertible.
	\end{definition}

	\begin{corollary}
		Let $\ideal{p}$ be a prime ideal in some ring $R$. Recall that by definition the complement $S = R\setminus \ideal{p}$ is a multiplicative system. The localization $R_\ideal{p} = S^{-1}R$ is a local ring with maximal ideal $\ideal{m}_{R_\ideal{p}} = S^{-1}\ideal{p}$.
	\end{corollary}

	\newpage
	\subsection{Factorial Rings and Principal Ideal Domains}

	\begin{definition}
		Let $R$ be a ring. An element $a$ \textbf{divides} another element $r$ (or is a \textbf{divisor} of $r$), if there is an element $b$ such that $ab = r$. In this case we write $a \divides r$.

		Two elements $a$ and $b$ are \textbf{associate}, if there is a unit $u$ such that $au = b$. Being associate defines an equivalence relation on $R$ and we denote the equivalence class corresponding to $a$ by $aR^\times$. 
	\end{definition}

	\begin{lemma}[Properties of Divisibility]
		Given a ring $R$ the following statements hold.
		\begin{enumerate}[(i)]
			\item{
				If $a \divides r$ and $a$ is a zero divisor, then $r$ is a zero divisor.\\
				If $a \divides r$ and $a$ is not a zero divisor, then the corresponding factor $b$ with $ab = r$ is unique.
			}
			\item{
				\TODO{all zero divisors are associate? one zero divisor then all associate zero divisors?}
			}
			\item{
				If $R$ is an integral domain, then divisibility defines a \TODO{partial?} order on the set $R/R^\times$.
			}
		\end{enumerate}
	\end{lemma}

	\begin{definition}
		Let $R$ be a commutative ring.

		An element $r$ is \textbf{irreducible}, if it is not a unit and $ab = r$ implies that either $a$ or $b$ are a unit. Note that the first condition makes this an exclusive or and the second condition can be reformulated as any divisor is associate to $r$.

		An element $p$ is \textbf{prime}, if it is nonzero, not a unit and for any two elements $a,b$ the statement $p \divides ab$ implies $p \divides a$ or $p \divides b$.
	\end{definition}

	\begin{lemma}[Properties of Primes and Irreducibles]
		Let $R$ be a ring. The following statements hold.
		\begin{enumerate}[(i)]
			\item{
				Irreducible elements are not zero divisors and are in particular nonzero.
			}
			\item{
				In an integral domain any prime element is irreducible
			}
			\item{
				In an integral domain prime elements $p,q$ with $p \divides q$ are associate.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				If $R$ is trivial $0$ is a unit.
			}
			\item{
				Use the cancellation rule of integral domains.
			}
			\item{
				$ap=q$ implies $qb=a$ or $qc=p$. Applying the cancellation rule on $q=ap=qbp$ yields a contradiction to $p$ not being a unit. Hence the cancellation rule on $q=ap=qac$ shows that $p$ and $q$ are associate via the unit $a$.
			}
		\end{enumerate}
	\end{sketch}

	\begin{definition}
		An integral domain $R$ is a \textbf{unique factorization domain} (or simply \textbf{factorial}), if it satisfies one of the following equivalent conditions
		\begin{enumerate}[(i)]
			\item{
				Every nonzero element $r$ admits a factorization
				\begin{equation*}
					r = u\cdot q_1\cdot ...\cdot q_n
				\end{equation*}
				into a unit $u$ and (not necessarily distinct) irreducible elements $q_1,...,q_n$, which is unique in the sense that any such factorization has the same amount of factors and the irreducible elements are unique up to permutation and multiplication by units.
			}
			\item{
				In $R$ every irreducible element is prime and for every nonzero element $r$ there is a factorization
				\begin{equation*}
					r = u\cdot q_1\cdot ...\cdot q_n
				\end{equation*}
				into a unit $u$ and (not necessarily distinct) irreducible elements $q_1,...,q_n$.
			}
			\item{
				Every nonzero element $r$ admits a factorization
				\begin{equation*}
					r = u\cdot p_1\cdot ... \cdot p_n
				\end{equation*}
				into a unit $u$ and (not necessarily distinct) prime elements $p_1,...,p_n$.
			}
			\item{
				In $R$ every irreducible element is prime and ascending chains of principal ideals stabilize, ie. for
				\begin{equation*}
					\Gen{r_1} \subseteq \Gen{r_2} \subseteq ...
				\end{equation*}
				there is some $N \in \N$ such that for $n \geq N$ we find $\Gen{r_n} = \Gen{r_N}$.
			}
			\item{
				Every nonzero prime ideal contains a prime element (Kaplansky)
			}
		\end{enumerate}
	\end{definition}
	\begin{sketch}
		(i)$\implies$(ii) Use existence and uniqueness of factorizations to deduce prime element property

		(ii)$\implies$(iii) $\checkmark$

		(iii)$\implies$(i) Since in an integral domain every prime element is irreducible, it suffices to show uniqueness of the factorization. We can restrict ourselves to the comparison of two factorizations with primes $p_i$ and irreducibles $q_j$
		\begin{equation*}
			u\cdot p_1\cdot ...\cdot p_m = v \cdot q_1 \cdot ... \cdot q_n
		\end{equation*}
		By definition of primality we find up to reordering $p_1 \divides q_1$, ie. $q_1 = v_1q_1$ and by definition of irreducability $v_1 \in R^\times$, so $p_1$ and $q_1$ are associate. Using the cancellation rule for integral domains yields
		\begin{equation*}
			u\cdot p_2\cdot ...\cdot p_m = vv_1 \cdot q_2 \cdot ... \cdot q_n.
		\end{equation*}
		By induction (and contradiction to deduce $m=n$) we get the desired uniqueness property.
	\end{sketch}

	\begin{lemma}[Properties of Factorial Rings]
		\TODO{irreducibles are prime}

		\TODO{weird valuation https://math.stackexchange.com/a/2191363/598157}
	\end{lemma}

	\begin{definition}
		Let $R$ be a ring. A \textbf{principal ideal} is an ideal, which is generated by one element. A \textbf{principal ideal domain} is an integral domain, in which every ideal is a principal ideal.
	\end{definition}

	\TODO{R PID then prime ideal != 0 is maximal ideal}

	\begin{theorem}[Fundamental Theorem of Arithmetic]
		Let $R$ be a principal ideal domain. Then $R$ is a factorial ring, i.e. every element has a (in some sense) unique decomposition into irreducible elements.
	\end{theorem}

	\newpage
	\subsection{Noetherian Rings}

	\begin{definition}
		A ring $R$ is \textbf{noetherian}, if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				Every ascending chain $\ideal{a}_0 \diveq \ideal{a}_1 \diveq ...$ of ideals stabilizes, i.e. satisfies $\ideal{a}_n = \ideal{a}_N$ for some $N$ and all $n \geq N$.
			}
			\item{
				Every nonempty set of ideals has a maximal element with respect to inclusion.
			}
			\item{
				Every ideal $\ideal{a} \diveq R$ is finitely generated, i.e. $\ideal{a} = \Gen{a_1,...,a_n}$ for some $a_1,...,a_n \in R$.
			}
			\item{
				Every prime ideal $\ideal{p} \divneq R$ is finitely generated.
			}
		\end{enumerate}
	\end{definition}
	\begin{sketch}
		(i)$<==>$(ii) via Zorn\\
		(i)$<==>$(iii) by taking contrapositions\\
		(iii)$<==>$(iv) show that finitely generated proper ideals form an Ako family.
	\end{sketch}

	\begin{lemma}[Properties of Noetherian Rings]\vspace{-1.5em}
		\begin{enumerate}[(i)]
			\item{
				Factor rings and localizations of noetherian rings are noetherian.
			}
			\item{
				Let $R$ be a noetherian ring and $\ideal{a}$ be an ideal. The set of prime ideals containing $\ideal{a}$ has finitely many minimal elements. In particular $R$ has only finitely many minimal prime ideals.
			}
			\item{
				In a noetherian ring every descending chain $\ideal{p}_0 \videq \ideal{p}_1 \videq ...$ of prime ideals stabilizes.
			}
			\item{
				Noetherian rings are atomic, i.e. any element in a noetherian ring can be written as a product of irreducible elements.
			}
			\item{
				Every surjective endomorphism of a noetherian ring is an isomorphism.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				Use the correspondence theorems.
			}
			\item{
				Let $\mathcal{F}$ be the set of prime ideals containing $\ideal{a}$.

				\textit{Assume} there exits a countable subset $\mathcal{F}_0$ of the set of minimal elements of $\mathcal{F}$.
				\begin{tab}[1.3cm]
					Fix an indexing $(\ideal{p}_n)_n$ of $\mathcal{F}_0$ and let $\mathcal{F}_n = \mathcal{F}_0 \setminus \{\ideal{p}_1,...,\ideal{p}_n\}$. As $R$ is noetherian, the ascending sequence
					\begin{equation*}
						\Intersection \limits_{\ideal{p} \in \mathcal{F}_0} \ideal{p} \divneq \Intersection \limits_{\ideal{p} \in \mathcal{F}_1} \ideal{p} \divneq \Intersection \limits_{\ideal{p} \in \mathcal{F}_2} \ideal{p} \divneq ...
					\end{equation*}
					stabilizes for some $N \in \N$. But now
					\begin{equation*}
						\Intersection \limits_{\ideal{p} \in \mathcal{F}_{N+1}} \ideal{p} = \Intersection \limits_{\ideal{p} \in \mathcal{F}_{N}} \ideal{p} \subseteq \ideal{p}_{N+1},
					\end{equation*}
					\TODO{thus $\ideal{p}_m \subseteq \ideal{p}_{N+1}$} for some $m > N+1$. This contradicts minimality of $\ideal{p}_{N+1}$.
				\end{tab}
			}
			\item{

			}
			\item{

			}
			\item{
				\textit{Assume} there is a surjective but not injective endomorphism $f:R-->>R$.
				\begin{tab}[1.3cm]
					Then $\ker f \divneq \ker (f\circ f) \divneq ...$ is a nonstabilizing chain. A contradiction.
				\end{tab}
			}\vspace{-1.8em}
		\end{enumerate}
	\end{sketch}

	\newpage
	\subsection{Artinian Rings}

	\begin{definition}
		A ring $R$ is \textbf{artinian}, if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				Every descending chain $\ideal{a}_0 \videq \ideal{a}_1 \videq ...$ of ideals stabilizes.
			}
			\item{
				Every nonempty set of ideals has a minimal element with respect to inclusion.
			}
		\end{enumerate}
	\end{definition}

	\begin{lemma}[Properties of Artinian Rings]\vspace{-1.5em}
		\begin{enumerate}[(i)]
			\item{
				Factor rings and localizations of artinian rings are artinian.
			}
			\item{
				Every element is nilpotent or a unit.
			}
			\item{
				An artinian ring is an integral domain, if and only if it is a field. In particular every prime ideal is maximal.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{enumerate}[(i)]
			\item{
				Use the correspondence theorems.
			}
			\item{
				\textit{Assume} $r$ is neither nilpotent nor a unit.
				\begin{tab}[1.3cm]
					The chain $\Gen{r} \vidneq \Gen{r^2} \vidneq ...$ does not stabilize. A contradiction.
				\end{tab}
			}
			\item{
				By (i) and (ii).
			}\vspace{-1.8em}
		\end{enumerate}
	\end{sketch}

	\TODO{artinian vs noetherian}

	\subsection{Dedekind Domains?}


	\newpage
	\section{Algebras}
	\subsection{Algebras}

	\begin{definition}
		Let $R$ be a ring. A \textbf{$R$-algebra} is a ring $A$ together with a morphism of rings $\varphi_A:R-->A$ called the \textbf{structure morphism}. In more explicit terms $A$ is a ring  together with a scalar multiplication $\cdot: R \times A --> A$, such that the multiplication $\cdot: A \times A --> A$ is $R$-bilinear, i.e. satisfies
		\begin{equation*}
			 (r\cdot a)\cdot b = r\cdot(a\cdot b) = a \cdot (r \cdot b).
		\end{equation*}
		The structure morphism can be recovered from this description by considering
		\begin{equation*}
			\begin{array}{rcl}
				R & \longrightarrow & A\\
				r & \longmapsto & r\cdot 1_A.
			\end{array}
		\end{equation*}
		\begin{minipage}{\textwidth-4cm}
			A morphism of $R$-algebras is a morphism of rings $f:A-->B$ compatible with the structure morphisms in the sense that the diagram on the right commutes. Explicitly $f$ has to satisfy the additional requirement that
			\begin{equation*}
				f(ra) = rf(a).
			\end{equation*}
		\end{minipage}
		\begin{minipage}{4cm}
			\begin{equation*}
				\begin{diagram}
					\threebytwo[narrow]
						{}{R}{}
						{A}{}{B}

					\arrow{n}{sw}{\varphi_A}[above left]
					\arrow{n}{se}{\varphi_B}[above right]
					\arrow{sw}{se}{f}[below]
				\end{diagram}
			\end{equation*}
		\end{minipage}
		By construction the category of $R$-algebras, denoted $\ncat{Alg}_R$, is the comma category $R \downarrow \ncat{CRing}$.
	\end{definition}

	\subsection{Polynomial Rings}

	\begin{lemma}
		A polynomial ring $R[X]$ is a principal ideal domain, \TODO{if and only if} $R$ is a field.
	\end{lemma}

	\begin{theorem}[Stability]
		Let $R$ be a commutative ring.
		\begin{enumerate}[(i)]
			\item{
				If $R$ is an integral domain, so is $R[X]$.
			}
			\item{
				If $R$ is a factorial domain, so is $R[X]$.
			}
			\item{
				If $R$ is noetherian, so is $R[X]$ (Hilbert's Basis Theorem)
			}
			\item{
				\TODO{If $R$ is artinian, so is $R[X]$. (wrong?)}
			}
		\end{enumerate}
	\end{theorem}

	\subsection{$K$-algebras and Krull-Dimension}

	\begin{theorem}
		Let $K$ be a field. Then the Krull-dimension of $K[X_1,...,X_n]$ is $n$.
	\end{theorem}

	\subsection{Matrix Rings}

	\TODO{TODO matrix rings as algebras over $R$}
\end{document}