\input{../header.tex}

\title{
	$\Sum \text{math}(y)$\\
	Abelian Categories
}
\author{Jonas Linßen}

\begin{document}
	\maketitle
	\tableofcontents

	\newpage
	\section{Abelian Categories}
	\subsection{Pointed Categories}

	\begin{definition}[Pointed Category]
		A category $\cat{C}$ is \textbf{pointed}, if it has a \textit{zero object}, i.e. an object $0$, which is both initial and terminal. Recall that for given objects $X,Y$ in $\cat{C}$ the unique morphism $0_{XY}:X-->0-->Y$ is called the \textit{zero morphism} from $X$ to $Y$.
	\end{definition}

	\begin{definition}[Kernel, Cokernel]
		Let $\cat{C}$ be a pointed category and $f:X-->Y$ be a morphism in $\cat{C}$.

		The \textit{equalizer} $\ker f := \eq (f,0_{XY})$ is called the \textbf{kernel} of $f$, if it exists.\\
		Dually, the \textit{coequalizer} $\coker f := \coeq(f,0_{XY})$ is called \textbf{cokernel} of $f$, if it exists.
	\end{definition}

	\begin{lemma}[Properties of Kernels and Cokernels]
		Let $\cat{C}$ be a pointed category. Then
		\begin{enumerate}[(i)]
			\item{
				The kernel of $0:X-->Y$ is given by $1_X$ (up to isomorphism).\\
				Dually, the cokernel of $0:X-->Y$ is given by $1_Y$ (up to isomorphism)
			}
			\item{
				For ever monomorphism $f:X\longrightmono Y$ is holds that $\ker f = 0_{0X}$.\\
				Dually, for every epimorphism $f:X \longrightepi Y$ it holds that $\coker f = 0_{Y0}$
			}
			\item{
				For any morphisms $f:X-->Y$ and monomorphism $g:Y \longrightmono Z$ we have $\ker gf = \ker f$.\\
				Dually, for any epmorphisms $f:X \longrightepi Y$ and morphism $g:Y \longrightmono Z$ we have the identity $\coker gf = \coker g$.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		\begin{minipage}{\textwidth}
		\begin{center}
		\begin{tabular}{l|l}
			\makecell{
				(i) Consider\\
				\begin{diagram}
					\threebytwo
						{}{T}{}
						{X}{X}{Y}
					\arrow[dashed]{n}{sw}{}
					\arrow{n}{s}{t}[right]
					\arrow{sw}{s}{1_X}[below]
					\arrow[higher]{s}{se}{0_{XY}}[above]
					\arrow[lower]{s}{se}{0_{XY}}[below]
				\end{diagram}
			}
			&
			\makecell{
				(ii) Consider\\
				\begin{diagram}
					\threebytwo
						{}{T}{}
						{0}{X}{Y}
					\arrow[dashed]{n}{sw}{}
					\arrow{n}{s}{t}[right]
					\arrow{sw}{s}{0_{0X}}[below]
					\arrow[higher,mono]{s}{se}{f}[above]
					\arrow[lower]{s}{se}{0_{XY}}[below]
				\end{diagram}\vspace{1em}\\
				and calculate $ft = 0t = 0 = f0$.
			}\\
			\hline
			\makecell{
				(iii) Consider\\
				\begin{diagram}
					\threebytwo
						{}{T}{}
						{K}{X}{Z}
					\arrow[dashed]{n}{sw}{}
					\arrow{n}{s}{t}[right]
					\arrow{sw}{s}{\ker gf}[below]
					\arrow[higher]{s}{se}{gf}[above]
					\arrow[lower]{s}{se}{0_{XZ}}[below]
				\end{diagram}\vspace{1em}\\
				and observe that\\
				$gft = 0t = g0t$ implies $ft = 0t$.
			}
		\end{tabular}
		\end{center}
		\end{minipage}
	\end{sketch}

	Note that a kernel is always mono and a cokernel is always epi. This leads to the following definition.

	\begin{definition}[Normal Mono, Normal Epi]
		Let $\cat{C}$ be a pointed category.

		A monomorphism is \textbf{normal}, if it is a kernel.\\
		Dually, an epimorphism is \textbf{normal}, if it is a cokernel.
	\end{definition}

	\begin{lemma}[Characterization of Normal Morphisms]
		Let $\cat{C}$ be a pointed category, $f:X-->Y$ be a morphism in $\cat{C}$ and assume that the following kernels and cokernels exist.

		Then $f$ is a normal monomorphism, if and only if $f = \ker \coker f$.\\
		Dually, $f$ is a normal epimorphism, if and only if $f = \coker \ker f$.
	\end{lemma}
	\begin{proof}
		\enquote{$<==$}:\\
		Holds by definition.

		\enquote{$==>$}:\\
		Let $t:T-->Y$ with \textcolor{blue}{$ct = 0t = 0$}. Since \textcolor{orange}{$gf=0f$} $=0=$ \textcolor{darkgreen}{$g0=gf$} there is a unique $z:C-->Z$ such that $zc = g$. Now \textcolor{blue}{$zct = z0t$} $=$ \textcolor{orange}{$0t=gt$} $=zct$, hence there is a unique $x:T-->X$ satisfying $fx=t$.
		\begin{equation*}
			\begin{matrix}
			\begin{diagram}
				\threebythree
					{}{T}{}
					{X}{Y}{C}[gray]
					{}{Z}{}

				\arrow[dashed,orange]{n}{w}{\exists!x}[above left]
				\arrow[orange]{n}{c}{t}[right]

				\arrow[higher,orange]{w}{c}{f}[above]
				\arrow[lower,gray]{w}{c}{0}[below]
				\arrow[higher,gray]{c}{e}{c}[above]
				\arrow[lower,gray]{c}{e}{0}[below]

				\arrow[lefter,orange]{c}{s}{0}[left]
				\arrow[righter,orange]{c}{s}{g}[right]
				\arrow[dashed,gray]{e}{s}{z}[below right]
			\end{diagram}&
			\begin{diagram}
				\threebythree
					{}{T}[gray]{}
					{X}{Y}{C}
					{}{Z}{}

				\arrow[dashed,gray]{n}{w}{x}[above left]
				\arrow[gray]{n}{c}{t}[right]

				\arrow[higher,darkgreen]{w}{c}{f}[above]
				\arrow[lower,darkgreen]{w}{c}{0}[below]
				\arrow[higher,darkgreen]{c}{e}{c}[above]
				\arrow[lower,gray]{c}{e}{0}[below]

				\arrow[lefter,gray]{c}{s}{0}[left]
				\arrow[righter,darkgreen]{c}{s}{g}[right]
				\arrow[dashed,darkgreen]{e}{s}{\exists!z}[below right]
			\end{diagram}&
			\begin{diagram}
				\threebythree
					{}{T}{}
					{X}{Y}{C}
					{}{Z}[gray]{}

				\arrow[dashed,blue]{n}{w}{\exists!x}[above left]
				\arrow[blue]{n}{c}{t}[right]

				\arrow[higher,blue]{w}{c}{f}[above]
				\arrow[lower,gray]{w}{c}{0}[below]
				\arrow[higher,blue]{c}{e}{c}[above]
				\arrow[lower,blue]{c}{e}{0}[below]

				\arrow[lefter,gray]{c}{s}{0}[left]
				\arrow[righter,gray]{c}{s}{g}[right]
				\arrow[dashed,gray]{e}{s}{z}[below right]
			\end{diagram}\\
			{\color{orange}f\text{ normal mono}}
			&{\color{darkgreen}c=\coker f}
			&{\color{blue}f = \ker c = \ker \coker f}
			\end{matrix}\vspace{-2em}
		\end{equation*}
	\end{proof}

	\newpage
	\subsection{Ab-enriched, Additive and Abelian Categories}

	\begin{definition}[$\ncat{Ab}$--enriched Category]
		A locally small category $\cat{A}$ is \textbf{$\ncat{Ab}$-enriched}, if every homset $\cat{A}(A,B)$ is an abelian group and the composition $\circ_{ABC}: \cat{A}(B,C) \times \cat{A}(A,B) --> \cat{A}(A,C)$ is a group homomorphism in both variables.
	\end{definition}

	\begin{definition}[Biproduct]
		Let $\cat{A}$ be an $\ncat{Ab}$-enriched category and $A,B$ be two objects. The \textbf{biproduct} $A \oplus B$ of $A$ and $B$ is an object $C$, which satisfies one of the following equivalent conditions.

		\begin{enumerate}[(i)]
			\item{
				There are morphisms
				\begin{equation*}
					\begin{diagram}
						\threebyone
							{A}{C}{B}
						\arrow[higher]{c}{w}{\pi_A}[above]
						\arrow[lower]{w}{c}{\iota_A}[below]
						\arrow[higher]{c}{e}{\pi_B}[above]
						\arrow[lower]{e}{c}{\iota_A}[below]
					\end{diagram}
				\end{equation*}
				% $$
				% 	\begin{matrix}
				% 		\pi_A:C-->A,&\iota_A:A-->C,\\
				% 		\pi_B:C-->B,&\iota_B:B-->C,
				% 	\end{matrix}
				% $$
				which satisfy the equations
				\begin{align*}
					\pi_A\iota_A, = 1_A, \hspace{1cm}\pi_B\iota_A = 0,\\
					\pi_B\iota_B, = 1_B, \hspace{1cm}\pi_A\iota_B = 0,\\
					\iota_A\pi_A + \iota_B\pi_B = 1_C\hspace{.7cm}
				\end{align*}
			}
			\item{
				$C$ is the product $A \times B$.
			}
			\item{
				$C$ is the coproduct $A + B$.
			}
		\end{enumerate}
	\end{definition}
	We have to show that (i) - (iii) are indeed equivalent
	\begin{sketch}
		\enquote{(i) $==>$ (ii)}:\\
		For $t_A:T-->A,t_B:T-->B$ note that $t := \iota_At_A + \iota_Bt_B$ satisfies the universal property of the product.

		\enquote{(ii) $==>$ (i)}:\\
		Use the first four equations to define $\iota_A$ and $\iota_B$ via the universal property. Observe that $\pi_A(\iota_A\pi_A + \iota_B\pi_B) = \pi_A$ and $\pi_B(\iota_A\pi_A + \iota_B\pi_B) = \pi_B$ and use the universal property to derive the fifth equation.

		\enquote{(i) $<==>$ (iii)}: By duality.
	\end{sketch}

	\begin{definition}[Additive category, Abelian category]
		A locally small category $\cat{A}$ is \textbf{additive}, if
		\begin{enumerate}[$\circ$]
			\item{it is pointed, i.e. has a zero object,}
			\item{is $\ncat{Ab}$-enriched}
			\item{and has all binary biproducts.}
		\end{enumerate}

		An \textbf{abelian} category is an additive category, which
		\begin{enumerate}[$\circ$]
			\item{has all kernels and cokernels}
			\item{and for which ever mono- or epimorphism is normal.}
		\end{enumerate}
	\end{definition}

	\begin{proposition}[Characterization of Equalizers and Coequalizers]
		Let $\cat{A}$ be an additive category and $f_1,f_2$ be parallel morphisms.

		Then, if the equalizer $\eq(f_1,f_2)$, the kernel $\ker(f_1-f_2)$ or the kernel $\ker(f_2-f_1)$ exists, so do the others and they are isomorphic.\\
		Dually, if the coequalizer $\coeq(f_1,f_2)$, the cokernel $\coker(f_1-f_2)$ or the cokernel $\coker(f_2-f_1)$ exists, so do the others and they are isomorphic.
	\end{proposition}
	\begin{sketch}
		TODO
	\end{sketch}

	\begin{corollary}[(Co-)Completeness of Abelian Categories]
		Any abelian category is complete and cocomplete.
	\end{corollary}
	\begin{sketch}
		TODO
	\end{sketch}

	\begin{theorem}[Uniqueness of Additive Structure]
		The $\ncat{Ab}$-enrichment of an abelian group is unique (up to group isomorphism).
	\end{theorem}
	\begin{sketch}
		TODO
	\end{sketch}

	\begin{proposition}[Characterization of Monos, Epis and Isos]
		Let $\cat{A}$ be an abelian category and $f:A-->B$ be a morphism.
		
		~\\
		Then the following are equivalent.
		\begin{enumerate}[(i)]
			\item{
				The morphism $f$ is a monomorphism.
			}
			\item{
				The kernel of $f$ satisfies $\ker f = 0$.
			}
			\item{
				For any morphism $t:T-->A$ it holds that $ft = 0$ implies $t = 0$.
			}
		\end{enumerate}
		Dually, the following are equivalent.
		\begin{enumerate}[(i)]
			\item{
				The morphism $f$ is an epimorphism.
			}
			\item{
				The cokernel of $f$ satisfies $\coker f = 0$.
			}
			\item{
				For any morphism $t:B-->T$ it holds that $tf = 0$ implies $t = 0$.
			}
		\end{enumerate}

		~\\
		Moreover it holds that $f$ is an isomorphism, if and only if it is both mono and epi.
	\end{proposition}
	\begin{sketch}
		TODO
	\end{sketch}

	\begin{corollary}[Characterization of Abelian Categories]
		A locally small category $\cat{A}$ is an abelian category if and only if it is a complete and cocomplete category with finite biproducts, where every monomorphism and every epimorphism is normal.
	\end{corollary}

	\newpage
	\subsection{Images, Kernels and Cokernels}

	\begin{theorem}[Epi-Mono Factorization]
		Let $\cat{A}$ be an abelian category. Every morphism $f:A-->B$ admits a factorization $f=ip$, where $i = \ker \coker f$ and $p = \coker \ker f$, which is unique (up to isomorphism).
	\end{theorem}
	\begin{sketch}
		TODO
	\end{sketch}

	\begin{definition}[Image]
		Let $\cat{C}$ be a pointed category and $f:X-->Y$ be a morphism.

		The \textbf{image} of $f$ is defined to be $\im f := \ker \coker f$, if the corresponding kernel and cokernel exist.
	\end{definition}

	\begin{theorem}[Isomorphism Theorem]
		TODO
	\end{theorem}

	\begin{definition}[Quotient]
		TODO
	\end{definition}

	\TODO{pullbacks and pushouts?}

	\newpage
	\subsection{Additive Functors}

	\begin{definition}[Additive Functor]
		A functor $\fun{F}:\cat{A} --> \cat{B}$ between additive categories is \textbf{additive}, if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				For any pair of objects $A,A'$ in $\cat{A}$ the map 
				\begin{equation*}
					\begin{array}{rcl}
						F_{AA'}: \cat{A}(A,A') &\longrightarrow& \cat{B}(\fun{F}A, \fun{F}A')\\
						f & \longmapsto& \fun{F}f
					\end{array}
				\end{equation*}
				is a group homomorphism.
			}
			\item{
				$\fun{F}$ preserves biproducts / products / coproducts
			}
		\end{enumerate}
	\end{definition}

	\begin{lemma}[Representables Are Additive]
		TODO
	\end{lemma}

	\begin{lemma}[$\Add(A,B)$ is $\ncat{Ab}$-enriched]
		TODO
	\end{lemma}

	\begin{theorem}[Additive Yoneda Lemma]
		TODO
	\end{theorem}

	\begin{proposition}[Characterization of Additive Functors]
		
	\end{proposition}

	\newpage
	\subsection{Exact Sequences, Diagram Lemmas and Exact Functors}

	\begin{definition}
		Let $\cat{A}$ be an abelian category. A sequence $A \xrightarrow{f} B \xrightarrow{g} C$ is \textbf{exact in $B$}, if $\im f \isom \ker g$. An arbitrary sequence
		\begin{align*}
			... --> A_{n+1} --> A_n --> A_{n-1} --> ...
		\end{align*}
		is \textbf{exact}, if it is exact in every (interior) $A_i$.

		A \textbf{short exact sequence} is an exact sequence of the form
		\begin{align*}
			0 --> A --> B --> C --> 0
		\end{align*}

		Note that this definitions are again selfdual.
	\end{definition}

	\begin{lemma}[Properties of Exact Sequences]
		Let $\cat{A}$ be an abelian category.
		\begin{enumerate}[(i)]
			\item{
				The sequence $0 --> A \xrightarrow{f} B$ is exact, if and only if $f$ is mono.

				Dually, the sequence $B \xrightarrow{g} C --> 0$ is exact, if and only if $g$ is epi.

				In particular, the sequence $0 --> A \xrightarrow{f} B --> 0$ is exact, if and only if $f$ is an isomorphism.
			}
			\item{
				The sequence $0 --> A \xrightarrow{f} B \xrightarrow{g} C$ is exact, if and only if $f \isom \ker g$.

				Dually, the sequence $A \xrightarrow{f} B \xrightarrow{g} C --> 0$ is exact, if and only if $g \isom \coker f$.
			}
			\item{
				The sequence $A \xrightarrow{f} B \xrightarrow{g} C$ is exact, if and only if the sequence $0 --> I \xrightarrow{i} B \xrightarrow{q} J --> 0$ is a short exact sequence, where $f = ip$, $g=jq$ are image factorizations.
			}
		\end{enumerate}
	\end{lemma}
	\begin{sketch}
		For (i) recall that $f$ is mono  iff $\ker f = 0 \;(= \im f)$.\\
		For (ii) note \TODO{todo}
		% \begin{enumerate}[(i)]
		% 	\item{
				
		% 	}
		% \end{enumerate}
	\end{sketch}

	\begin{definition}[Split Exact Sequence]
		Let $\cat{A}$ be an abelian category. A short exact sequence $0 --> A \xrightarrow{f} B \xrightarrow{g} C --> 0$ is said to be \textbf{split exact} if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				$g$ is split epi i.e. has a section $s:C --> B$
			}
			\item{
				$f$ is split mono i.e. has a retraction $r:B --> A$.
			}
			\item{
				There are morphisms $s:C --> B$ and $r:B-->A$, which turn
				\begin{equation*}
					\begin{diagram}
						\threebyone
							{A}{B}{C}

						\arrow[lower]{w}{c}{f}[below]
						\arrow[higher]{c}{e}{g}[above]
						\arrow[lower]{e}{c}{s}[below]
						\arrow[higher]{c}{w}{r}[above]
					\end{diagram}
				\end{equation*}
				into a biproduct $B \isom A \oplus B$.
			}
		\end{enumerate}
	\end{definition}
	\begin{sketch}
		(iii)$==>$(i)+(ii) by definition.\\
		(i)$==>$(iii) Consider $1_B - sg: B --> B$ and note that it satisfies $g(1_B - sg) = 0$. Since $f \isom \ker g$ there is a unique $r$ \TODO{tofu}
	\end{sketch}

	\begin{theorem}[Snake Lemma]
		\TODO{todo}
	\end{theorem}

	\begin{corollary}[Five Lemma]
		\TODO{todo}
	\end{corollary}

	\begin{corollary}[Niner Lemma]
		\TODO{todo}
	\end{corollary}

	\begin{definition}[Exact Functor]
		A functor $\fun{F}: \cat{A} --> \cat{B}$ between abelian categories is \textbf{left exact}, if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				$\fun{F}$ preserves exact sequences of the form $0 --> A --> B --> C$.
			}
			\item{
				$\fun{F}$ is additive and preserves kernels.
			}
			\item{
				$\fun{F}$ preserves finite limits.
			}
		\end{enumerate}

		Dually, $\fun{F}$ is called \textbf{right exact} if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				$\fun{F}$ preserves exact sequences of the form $A --> B --> C --> 0$.
			}
			\item{
				$\fun{F}$ is additive and preserves cokernels.
			}
			\item{
				$\fun{F}$ preserves finite colimits.
			}
		\end{enumerate}

		Finally, $\fun{F}$ is \textbf{exact}, if it satisfies one of the following equivalent conditions.
		\begin{enumerate}[(i)]
			\item{
				$\fun{F}$ is both left and right exact.
			}
			\item{
				$\fun{F}$ preserves short exact sequences.
			}
			\item{
				$\fun{F}$ preserves arbitrary exact sequences.
			}
		\end{enumerate}

		Note that in particular each left exact / right exact / exact functor is additive.
	\end{definition}

	\begin{lemma}
		Let $\fun{F}: \cat{A} --> \cat{B}$ be a faithful exact functor. Then $\fun{F}$ reflects (not necessarily short) exact sequences.

		Analogue statements hold for faithful left exact / right exact functors.
	\end{lemma}
	\begin{sketch}
		Recall that faithful functors reflect all limits and colimits, which they preserve.
	\end{sketch}

	\newpage
	\section{(Co-)Homology and Derived Functors}
	\subsection{Chain Complexes and Homology}

	\begin{definition}
		Let $\cat{A}$ be an abelian category. A \textbf{chain complex} $C_\bullet$ in $\cat{A}$ is an infinite sequence
		\begin{align*}
			... --> C_{n+1} \xrightarrow{\partial_{n+1}} C_{n} \xrightarrow{\partial_n} C_{n-1} --> ...
		\end{align*}
		such that $\partial_n \partial_{n+1} = 0$ for all $n \in \Z$. %The morphisms $\partial_\bullet$ are called boundary morphisms or differentials of the chain complex.

		A morphism $f_\bullet: C_\bullet --> D_\bullet$ of chain complexes is a sequence $(f_n: C_n --> B_n)_n$ such that for all $n$ the following diagram commutes.
		\begin{equation*}
			\begin{diagram}
				\fourbytwo
					{...}{C_{n+1}}{C_n}{...}
					{...}{D_{n+1}}{D_n}{...}

				\arrow{nww}{nw}{}
				\arrow{nw}{ne}{\partial_{n+1}}[above]
				\arrow{ne}{nee}{}

				\arrow{sww}{sw}{}
				\arrow{sw}{se}{\partial_{n+1}}[below]
				\arrow{se}{see}{}

				\arrow{nw}{sw}{f_{n+1}}[left]
				\arrow{ne}{se}{f_{n}}[right]
			\end{diagram}
		\end{equation*}

		This defines a category $\Ch(\cat{A})$ of chain complexes over $\cat{A}$.

		A chain complex $C_\bullet$ is \textbf{concentrated in nonnegative degree}, if $C_n = 0$ for all $n < 0$, and \textbf{concentrated in nonpositive degree}, if $C_n = 0$ for all $n > 0$. We write $\Ch(\cat{A})^{+}$ and $\Ch(\cat{A})^{-}$ for the corresponding full subcategories of $\Ch(\cat{A})$.
	\end{definition}

	% \begin{remark}
	% 	The chain complexes concentrated in nonpositive degree are sometimes called \textbf{cochain complexes}.
	% 	To avoid writing negative indices when working with 
	% \end{remark}

	\begin{lemma}[Properties of $\Ch(\cat{A})$]
		\TODO{$\Ch(\cat{A})$ is an abelian category}

		\TODO{Tensor product and internal hom}
	\end{lemma}

	\begin{lemma}
		\TODO{preparation for homology functor}
	\end{lemma}

	\begin{definition}
		Let $\cat{A}$ be an abelian category. The \textbf{$n$-th homology functor} is defined by the assignment
		\begin{equation*}
			\begin{array}{rcl}
				\Ch(\cat{A}) & \xrightarrow{H_n} & \cat{A}\\
				C_\bullet & \longmapsto & H_n(C_\bullet) := \ker \partial_n / \im \partial_{n+1}\\
				{[f:C_\bullet \rightarrow D_\bullet]} & \longmapsto & {[H_n(f):H_n(C_\bullet) \rightarrow H_n(D_\bullet)]}
			\end{array}
		\end{equation*}
	\end{definition}

	\newpage
	\subsection{Projective and Injective Objects}

	\begin{definition}
		Let $\cat{A}$ be an abelian category. 

		An object $P$ in $\cat{A}$ is \textbf{projective}, if it satisfies one of the following equivalent conditions.
		\vspace{.5em}
		\begin{enumerate}[(i)]
			\item{
				\begin{minipage}[t]{\linewidth-4cm}
					For all epimorphisms $T_1 -->> T_2$ and morphisms $P --> T$ there exists a (not necessarily unique) lift as depicted on the right.
				\end{minipage}
				\begin{minipage}[t]{4cm}
					\vspace{-2em}
					\begin{equation*}
						\begin{diagram}
							\twobytwo[small]
								{}{T_1}
								{P}{T_2}

							\arrow{sw}{se}{}
							\arrow[epi]{ne}{se}{}
							\arrow[dashed]{sw}{ne}{\exists}[above left]
						\end{diagram}
					\end{equation*}
				\end{minipage}
				\vspace{-3em}
			}
			\item{
				$\cat{A}(P,-): \cat{A} --> \ncat{Ab}$ is exact.
			}
			\item{
				Any epimorphism $T -->> P$ is a split epimorphism.
			}
		\end{enumerate}

		Dually, an object $I$ in $\cat{A}$ is \textbf{injective}, if it satisfies one of the following equivalent conditions.
		\vspace{.5em}
		\begin{enumerate}[(i)]
			\item{
				\begin{minipage}[t]{\linewidth-4cm}
					For all monomorphisms $T_1 >--> T_2$ and morphisms $T_1 --> I$ there exists a (not necessarily unique) extension as depicted on the right.
				\end{minipage}
				\begin{minipage}[t]{4cm}
					\vspace{-2em}
					\begin{equation*}
						\begin{diagram}
							\twobytwo[small]
								{T_1}{T_2}
								{I}{}

							\arrow{nw}{sw}{}
							\arrow[mono]{nw}{ne}{}
							\arrow[dashed]{sw}{ne}{\exists}[below right]
						\end{diagram}
					\end{equation*}
				\end{minipage}
				\vspace{-3em}
			}
			\item{
				$\cat{A}(-,I): \cat{A}^{op} --> \ncat{Ab}$ is exact.
			}
			\item{
				Any monomorphism $I >--> T$ is a split monomorphism.
			}
		\end{enumerate}
	\end{definition}

	\subsection{Derived Functors}

	\newpage
	\section{The Freyd Mitchel Embedding Theorem}

\end{document}